---
title: "Courtship_Code_Summary_Compilation"
author: "Nathan Phipps"
Last Edited: "2/1/2021"
output: word_document
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
knitr::opts_chunk$set(fig.width=8, fig.height=5) 
```
Loading all packages
```{r}
library(tidyverse)
library(lmerTest)
library(plyr)
library(dplyr)
library(car)
require(MASS)
library(mlmRev)
library(agridat)
library(MCMCglmm)
library(ggplot2)
library(emmeans)
```

### INITIAL IMPORT, INSPECTION, AND CLEANUP OF DATA

Importing and checking structure of data frame. CourtMCMC is created because I will alter the dataframe differently in order to run MCMC, which would disrupt further non-MCMC analysis.
```{r}
Court <- read.csv("C:/Users/phipp/OneDrive/Documents/Grad School/Courtship Analysis/Data/FINAL_court.csv")
CourtMCMC <- read.csv("C:/Users/phipp/OneDrive/Documents/Grad School/Courtship Analysis/Data/FINAL_court.csv")
str(Court)
```

Specifying data types and variable types for each column
```{r}
cols(ID = col_integer(),
     Pop = col_character(),
     Family = col_integer(),
     DateBorn = col_character(),
     Sex = col_character(),
     Length = col_double(),
     Stage = col_character(),
     Date = col_character(),
     Coder = col_character(),
     Chase = col_integer(),
     Forced_Cop_Attempt = col_integer(),
     G_Swing = col_integer(),
     Sigmoid = col_integer(),
     Suc_Sigmoid_Cop = col_integer(),
     Ignore = col_integer(),
     Lunge = col_integer(),
     Nip = col_integer(),
     Suc_Forced_Cop = col_integer(),
     Copulated = col_character(),
     TTSigmoid = col_integer(),
     TTAvgSigmoid = col_integer())

#Making four variables into factors
Court %>%
  mutate(Pop = as.factor(Pop)) %>%
  mutate(Stage = as.factor(Stage)) %>%
  mutate(Family = as.factor(Family)) %>%
  mutate(ID = as.factor(ID)) -> Court

```

Checking the sample size for each encounter with female and each genetic line
```{r}
#How many "first encounter" trials?
nrow(Court[Court$Stage == "Before",])
#How many "second encounter" trials?
nrow(Court[Court$Stage == "After",])
#Create dataframe object of ONLY "second encounter" trials
Court.tmp <- Court [Court$Stage == "After",]

#EXAMPLE: How many "second encounter" trials for the Campo line? (Can be edited for any genetic line/encounter combo)
length (unique(Court.tmp$Pop == "CM"))

#How many total trials for the Campo line? (Includes both trials for each ind.)
nrow(Court[Court$Pop == "CM",])
#Create dataframe object of all trials for ONLY Campo individuals
Court.tmp1 <- Court [Court$Pop == "CM",]
#How many Campo individuals were observed?
length (unique(Court.tmp1$ID))

nrow(Court [Court$Pop == "CQ",])
Court.tmp2 <- Court [Court$Pop == "CQ",]
#How many CQ intercross individuals were observed?
length (unique(Court.tmp2$ID))

nrow(Court[Court$Pop == "QC",])
Court.tmp3 <- Court [Court$Pop == "QC",]
#How many QC intercross individuals were observed?
length (unique(Court.tmp3$ID))

nrow(Court[Court$Pop == "QH",])
Court.tmp4 <- Court [Court$Pop == "QH",]
#How many Quare individuals were observed?
length (unique(Court.tmp4$ID))
```

Checking family size for a single given family, and then a for loop to do the same for all families.
```{r}
#How many families are there?
length(unique(Court$Family))
#How many individuals are in family 401?
nrow(Court[Court$Family == "401",])
#How many "second encounter" trials?
nrow(Court[Court$Stage == "After",])
#Create dataframe object of ONLY "second encounter" trials
Court.tmp <- Court [Court$Stage == "After",]

#Creating an object called FamilySize for the for loop to use. I am unsure why this is necessary, but it's the only way I've found to make the loop run.
FamilySize <- 0
  
#For each family from 401 to 438, find the number of individuals in that family. If there are zero members of the family, skip it and move to the next family. Print the output next to the family number.

for(i in 401:438) {
  FamilySize[i] <- nrow(Court[Court$Family == i,])
     {if (FamilySize[i]==0)
    next
  } 
  FamilySize[i] <- nrow(Court[Court$Family == i,])
  print(c(i, FamilySize[i]))
}
```

### FIRST ANALYSIS STAGE: DATA DIAGNOSTICS AND CORRELATION TESTING ###

# Dataset prep
```{r}
#Apply factor mutation to characters
data_factor <- Court %>%
    mutate_if(sapply(Court, is.character), as.factor)
       #Check that it worked
sapply(data_factor, class)
```

```{r}
 #Visualize data in histogram
par(mfrow = c(1,3))
hist(data_factor$Forced_Cop_Attempt)
hist(data_factor$Sigmoid)
hist(data_factor$G_Swing)
```

# Statistical tests for best distribution

Forced copulation attempts distribution
```{r}
data_factor$Forced_Cop_Attempt.t <- data_factor$Forced_Cop_Attempt + 1 

 #normal test
 qqp(data_factor$Forced_Cop_Attempt.t, "norm")
```
```{r}
 #lognormal
 qqp(data_factor$Forced_Cop_Attempt.t, "lnorm")
``` 
```{r}
#Negative binomial
nbinom <- fitdistr(data_factor$Forced_Cop_Attempt.t, "Negative Binomial")
qqp(data_factor$Forced_Cop_Attempt.t, "nbinom", size = nbinom$estimate[[1]], mu = nbinom$estimate[[2]])
```
```{r}  
 #poisson
 poisson <- fitdistr(data_factor$Forced_Cop_Attempt.t, "Poisson")
qqp(data_factor$Forced_Cop_Attempt.t, "pois", lambda = poisson$estimate)
```
```{r}
 #gamma
gamma <- fitdistr(data_factor$Forced_Cop_Attempt.t, "gamma")
qqp(data_factor$Forced_Cop_Attempt.t, "gamma", shape = gamma$estimate[[1]], rate = gamma$estimate[[2]]) 
``` 

Sigmoids distribution
```{r}
data_factor$Sigmoid.t <- data_factor$Sigmoid + 1
qqp(data_factor$Sigmoid.t, "norm")
```
```{r}
   #lognormal
qqp(data_factor$Sigmoid.t, "lnorm")
```
```{r}
   #negative binomial
nbinom <- fitdistr(data_factor$Sigmoid.t, "Negative Binomial")
qqp(data_factor$Sigmoid.t, "nbinom", size = nbinom$estimate[[1]], mu = nbinom$estimate[[2]])
```
```{r}
   #poisson
poisson <- fitdistr(data_factor$Sigmoid.t, "Poisson")
qqp(data_factor$Sigmoid.t, "pois", lambda = poisson$estimate)
```
```{r}
   #gamma
gamma <- fitdistr(data_factor$Sigmoid.t, "gamma")
qqp(data_factor$Sigmoid.t, "gamma", shape = gamma$estimate[[1]], rate = gamma$estimate[[2]])
```

Gonopoodium Swings distribution
```{r}
data_factor$G_Swing.t <- data_factor$G_Swing + 1   

  #normal
qqp(data_factor$G_Swing.t, "norm")
```

```{r}
  #lognorm
qqp(data_factor$G_Swing.t, "lnorm")
```
```{r}
  #Negative Binomial
nbinom <- fitdistr(data_factor$G_Swing.t, "Negative Binomial")
qqp(data_factor$G_Swing.t, "nbinom", size = nbinom$estimate[[1]], mu = nbinom$estimate[[2]])
```
```{r}
  #poisson
poisson <- fitdistr(data_factor$G_Swing.t, "Poisson")
qqp(data_factor$G_Swing.t, "pois", lambda = poisson$estimate)
```
```{r}
  #gamma
gamma <- fitdistr(data_factor$G_Swing.t, "gamma")
qqp(data_factor$G_Swing.t, "gamma", shape = gamma$estimate[[1]], rate = gamma$estimate[[2]])
```

Model comparisons, residual tests, and other fit analyses were performed using the "try" function, but they're too janky and interfere with the knit function. See the file "Relevent code for Laura v. 2. R" for this code.

### SECOND ANALYSIS STAGE: LINEAR MODELS AND TRYING DIFFERENT DISTRIBUTION FITS ###

#Creating linear mixed models for sigmoids, forced copulations, and gonopodium swings using genetic line and experience as predictors, accounting for body length

Sigmoids
```{r}
lmm.Sig <- lmer(Sigmoid ~ Pop*Stage + Length 
                + (1|Family/ID),
                data=Court)

anova(lmm.Sig)

ranova(lmm.Sig)
```

Forced Copulations
```{r}
lmm.Force <- lmer(Forced_Cop_Attempt ~ Pop*Stage + Length
                  + (1|Family/ID),
                  data=Court)

anova(lmm.Force)

ranova(lmm.Force)
```

Gonopodium Swings 
```{r}
lmm.Swing <- lmer(G_Swing ~ Pop*Stage + Length
                  + (1|Family/ID),
                  data=Court)

anova(lmm.Swing)

ranova(lmm.Swing)
```

Turn ANOVAS into objects
```{r}
anova.Sig<-anova(lmm.Sig)
anova.Force<-anova(lmm.Force)
anova.Swing<-anova(lmm.Swing)
```

Attempt pairwise comparisons using emmeans and Tukey's HSD (These aren't running. Notes from STAT 511 should show how to make these models properly)
```{r}
#emmeans(lmm.Sig)

#aov.Swing<-aov(G_Swing ~ Pop*Stage + Length, data=Court)
#TukeyHSD(aov.Swing, which, ordered = FALSE, conf.level = 0.95)
```

### THIRD ANALYSIS STAGE: MCMCglmm (Markov Chain Monte Carlo)

Changing column name from "Family" to "famil" to avoid confusion with MCMC function "family", and checking that it worked
```{r}
colnames(CourtMCMC)[colnames(CourtMCMC) == "Family"] <- "famil"
str(CourtMCMC)
```

Creating Gaussian regression models using genetic line and experience as interactive predictors
```{r}
fit_gaussian_Sig <- MCMCglmm(Sigmoid ~ Pop*Stage,
                         random = ~idv(1):famil + idv(1):ID,
                         data = CourtMCMC,
                         nitt=20000)
fit_gaussian_Cop <- MCMCglmm(Forced_Cop_Attempt ~ Pop*Stage,
                             random = ~idv(1):famil + idv(1):ID,
                             data = CourtMCMC,
                             nitt=20000)
fit_gaussian_Swing <- MCMCglmm(G_Swing ~ Pop*Stage,
                               random = ~idv(1):famil + idv(1):ID,
                               data = CourtMCMC,
                               nitt=20000)
```

Creating Poisson regression models using genetic line and experience as interactive predictors
```{r}
fit_poisson_Sig <- MCMCglmm(Sigmoid ~ Pop*Stage,
                            random = ~idv(1):famil + idv(1):ID,
                            family = "poisson", data = CourtMCMC,
                            nitt=20000)
fit_poisson_Cop <- MCMCglmm(Forced_Cop_Attempt ~ Pop*Stage,
                            random = ~idv(1):famil + idv(1):ID,
                            family = "poisson", data = CourtMCMC,
                            nitt=20000)
fit_poisson_Swing <- MCMCglmm(G_Swing ~ Pop*Stage,
                            random = ~idv(1):famil + idv(1):ID,
                            family = "poisson", data = CourtMCMC,
                            nitt=20000)

#Using glmer for regression as comparison
fit_glmer <- glmer(Sigmoid ~ Pop*Stage +
                     (1|famil/ID),
                   family = "poisson", data = CourtMCMC)
    #Bad things? I don't understand
```

Examining results
```{r}
summary(fit_gaussian_Sig)
summary(fit_gaussian_Cop)
summary(fit_gaussian_Swing)

str(fit_gaussian_Sig)

summary(fit_poisson_Sig)
summary(fit_poisson_Cop)
summary(fit_poisson_Swing)

summary(fit_glmer)

#Note: summary.MCMCglmm(fit_gaussian_Swing) returns the same results as the regular summary() function
```

Trying emmeans to find pairwise comparisons (Doesn't run yet. Use Stat 511 notes to construct the function)
```{r}
#emmeans(fit_gaussian_Cop)
```

### GENERATING FIGURES ###
Function for getting summary, created by Laura Stein.
Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
Arguments:
data: a data frame.
measurevar: the name of a column that contains the variable to be summariezed
groupvars: a vector containing names of columns that contain grouping variables
na.rm: a boolean that indicates whether to ignore NA's
conf.interval: the percent range of the confidence interval (default is 95%)
```{r}

summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
  library(plyr)
  
  # New version of length which can handle NA's: if na.rm==T, don't count them
  length2 <- function (x, na.rm=FALSE) {
    if (na.rm) sum(!is.na(x))
    else       length(x)
  }
  
  # This does the summary. For each group's data frame, return a vector with
  # N, mean, and sd
  datac <- ddply(data, groupvars, .drop=.drop,
                 .fun = function(xx, col) {
                   c(N    = length2(xx[[col]], na.rm=na.rm),
                     mean = mean   (xx[[col]], na.rm=na.rm),
                     sd   = sd     (xx[[col]], na.rm=na.rm)
                   )
                 },
                 measurevar
  )
  
  # Rename the "mean" column    
  datac <- rename(datac, c("mean" = measurevar))
  
  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
  
  # Confidence interval multiplier for standard error
  # Calculate t-statistic for confidence interval: 
  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
  datac$ci <- datac$se * ciMult
  
  return(datac)
}
```

Changing the position of a geom layer in ggplot, using the "dodge" function
```{r}
pd <- position_dodge(0.5) # move them 0.05 to the left and right

```

Line graph of Sigmoids vs Experience by Population, with standard error bars
```{r}
#Creating summary statistics table to use as basis for graph
dSig <- summarySE(Court, measurevar="Sigmoid",
                  groupvars=c("Pop","Stage"))

dSig

#Re-ordering levels before and after
dSig$Stage <- factor(dSig$Stage, levels = c("Before", "After"))

#Creating graph as object
graph.Sig <- ggplot(dSig, aes(x=Stage, y=Sigmoid, group=Pop, col=Pop)) +
  geom_errorbar(aes(ymin=Sigmoid-se, ymax=Sigmoid+se), position = pd, width=0.1, colour = "black", size=1) + 
  geom_line(size=1, position = pd) +
  geom_point(position = pd, size=3) +
  scale_fill_brewer(palette="Set1") +
  labs(x="Courtship Experience", y="# Sigmoids +/- SE") +
  scale_x_discrete(breaks=c("Before","After"),
                   labels=c("First","Second")) +
  guides(color=guide_legend("Genetic Line"))  +
  theme_classic() +
  expand_limits(y=c(0,14)) +
  theme(axis.title.x = element_text(face = "bold", colour = "black", size = 20, vjust = -1),
        axis.text.x  = element_text(vjust=0.5, size=18, face = "bold", colour = "black"),
        axis.title.y = element_text(face = "bold", colour = "black", size = 20, vjust = 3),
        axis.text.y = element_text(vjust = 0.5, size = 18, face = "bold", colour = "black"),
        legend.title = element_text(size = 13),
        legend.text = element_text(size = 11, face = "italic"),
        legend.position=c(0.9, 0.8),
        plot.margin=unit(c(1,1,1.5,1.2),"cm"))

graph.Sig
```

Doing the same thing for forced copulations
```{r}
#Creating summary statistics table to use as basis for graph
dForced <- summarySE(Court, measurevar="Forced_Cop_Attempt",
                     groupvars=c("Pop","Stage"))
# Re-ordering levels Before and After
dForced$Stage <- factor(dForced$Stage, levels = c("Before", "After"))

# Checking that reordering worked
head(dForced)
```

```{r}
#Creating graph as object
graph.Forced <- ggplot(dForced, aes(x=Stage, y=Forced_Cop_Attempt, group=Pop, col=Pop)) +
  geom_errorbar(aes(ymin=Forced_Cop_Attempt-se, ymax=Forced_Cop_Attempt+se), position = pd, width=0.1, colour="black", size=1) +
  geom_line(size=1, position = pd) + 
  geom_point(position = pd, size=3) +
  scale_fill_brewer(palette="Set1") +
  labs(x="Courtship Experience", y="# Forced Copulations +/- SE") +
  guides(color=guide_legend("Genetic Line"))  +
  scale_x_discrete(breaks=c("Before","After"),
                   labels=c("First","Second")) +
  expand_limits(y=c(0,14)) +
  theme_classic() +
  theme(axis.title.x = element_text(face = "bold", colour = "black", size = 20, vjust = -1),
        axis.text.x  = element_text(vjust=0.5, size=18, face = "bold", colour = "black"),
        axis.title.y = element_text(face = "bold", colour = "black", size = 20, vjust = 3),
        axis.text.y = element_text(vjust = 0.5, size = 18, face = "bold", colour = "black"),
        legend.title = element_text(size = 13),
        legend.text = element_text(size = 11, face = "italic"),
        legend.position=c(0.9, 0.8),
        legend.background = element_rect(size=0.5, color=1),
        plot.margin=unit(c(1,1,1.5,1.2),"cm"))

graph.Forced
```

Doing the same for gonopodium swings
```{r}
#Creating summary statistics table to use as basis for graph
dSwing <- summarySE(Court, measurevar="G_Swing",
                    groupvars=c("Pop","Stage"))

#Re-ordering levels before and after
dSwing$Stage <- factor(dSwing$Stage, levels = c("Before", "After"))


head(dSwing)
```
```{r}
graph.Swing <- ggplot(dSwing, aes(x=Stage, y=G_Swing, group=Pop, col=Pop)) +
  geom_errorbar(aes(ymin=G_Swing-se, ymax=G_Swing+se), position = pd, width=0.1, colour="black", size=1) +
  geom_line(size=1, position = pd) + 
  geom_point(position = pd, size=3) +
  scale_fill_brewer(palette="Set1") +
  labs(x="Courtship Experience", y="# Gonopodium Swings +/- SE") +
  guides(color=guide_legend("Genetic Line"))  +
  theme_classic() +
  scale_x_discrete(breaks=c("Before","After"),
                   labels=c("First","Second")) +
  expand_limits(y=c(0,14)) +
  theme(axis.title.x = element_text(face = "bold", colour = "black", size = 20, vjust = -1),
        axis.text.x  = element_text(vjust=0.5, size=18, face = "bold", colour = "black"),
        axis.title.y = element_text(face = "bold", colour = "black", size = 20, vjust = 3),
        axis.text.y = element_text(vjust = 0.5, size = 18, face = "bold", colour = "black"),
        legend.title = element_text(size = 13),
        legend.text = element_text(size = 11, face = "italic"),
        legend.position=c(0.15, 0.8),
        legend.background = element_rect(size=0.5, color=1),
        plot.margin=unit(c(1,1,1.5,1.2),"cm"))

graph.Swing
```

### BCFTools play
```{r}
pedigree <- read.csv("C:/Users/phipp/OneDrive/Documents/Grad School/Courtship Analysis/QTL_component/Guppy_pedigree_Mendelian_error_check.txt")

pedigree <- pedigree[,1:3] + 0.1

write.csv(pedigree,"Guppy_pedigree_Mendelian_error_check_2.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
```


